Transfer of Learned Opponent Models in Zero Sum Games

### 2019-12-17 Report/draft_report_v1.tex
added results section to report

---
title             : "Transfer of Learned Opponent Models in Zero Sum Games"
shorttitle        : "draft"

author: 
  - name          : "Ismail Guennouni"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Department of Experimental Psychology, University College London, 26 Bedford Way, London WC1H 0AP, United Kingdom"
    email         : "i.guennouni.17@ucl.ac.uk"
  - name          : "Maarten Speekenbrink"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Experimental Psychology, University College London"

author_note: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

@@

abstract: |
  Enter abstract here. Each new line herein must be indented, like this line.

keywords          : "keywords"
wordcount         : "X"

# bibliography      : ["r-references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
linkcolor         : "blue"
mask              : no
draft             : no

class             : "man"
output            : papaja::apa6_pdf
header-includes   :
  - \usepackage[document]{ragged2e}
---

```{r load-packages, include = FALSE}
library("papaja")
require(knitr)
require(citr)
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
```

```{r texreport}
texreport <- function(object,effect.name,...) UseMethod("texreport")
texreport.default <- function(object,effect.name,...) return(NULL)

texreport.summary.merMod <- function(object,effect.name,print.estimate=TRUE) {
  if(effect.name %in% rownames(object$coefficients)) {
    tab <- object$coefficients
    estimate <- tab[effect.name,"Estimate"]
    se <- tab[effect.name,"Std. Error"]
    Tval <- tab[effect.name,"t value"]
    df <- tab[effect.name,"df"]
    pval <- tab[effect.name,"Pr(>|t|)"]
    if(pval < .001) ptext <- "$p < .001$" else {
      if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
        ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
      }
    }
    txt <- paste("$b = ",round(estimate,2),"$, $SE = ",round(se,2),"$, $t(",round(df,2),") = ",round(Tval,2),"$, ",sep="")
    txt <- paste(txt,ptext,sep="")
  } else {
    txt <- "error"
  }
  txt
}

texreport.summary.Anova.mlm <- function(object,effect.name,effect.size=TRUE,sphericity.correction=TRUE) {
  if(effect.name %in% rownames(object$sphericity.tests) && sphericity.correction && object$sphericity.test[effect.name,"p-value"] < .05) {
    epsilon <- object$pval.adjustments[effect.name,"HF eps"]
    pval <- object$pval.adjustments[effect.name,"Pr(>F[HF])"]
  } else {
    epsilon <- 1
    pval <- object$univariate.tests[effect.name,"Pr(>F)"]
  }
  df1 <- object$univariate.tests[effect.name,"num Df"]*epsilon
  df2 <- object$univariate.tests[effect.name,"den Df"]*epsilon
  Fval <- object$univariate.tests[effect.name,"F"]
  etasq <- object$univariate.tests[effect.name,"SS"]/(object$univariate.tests[effect.name,"SS"] + object$univariate.tests[effect.name,"Error SS"])
  # generate latex output
  if(pval < .001) ptext <- "$p < .001$" else {
    if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
      ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
    }
  }
  txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
  if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
  txt <- paste(txt,ptext,sep="")
  txt
}

texreport.linearHypothesis.mlm <- function(object,effect.size=TRUE) {
  if(ncol(object$SSPH) == 1 & nrow(object$SSPH) == 1 & ncol(object$SSPE) == 1 & nrow(object$SSPE) == 1) {
    df1 <- object$df
    df2 <- object$df.residual
    Fval <- (object$SSPH/df1)/(object$SSPE/df2)
    pval <- 1-pf(Fval,object$df,object$df.residual)
    etasq <- object$SSPH/(object$SSPH + object$SSPE)
    # generate latex output
    if(pval < .001) ptext <- "$p < .001$" else {
      if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
        ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
      }
    }
    txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
    if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
    txt <- paste(txt,ptext,sep="")
  } else {
    txt <- "error"
  }
  txt
}

texreport.anova.lme <- function(object,effect.name,effect.size=FALSE) {
  if(effect.name %in% rownames(object)) {
    pval <- object[effect.name,"p-value"]
    if(is.null(pval)) pval <- object[effect.name,"Pr(>F)"]
    df1 <- object[effect.name,"numDF"]
    df2 <- object[effect.name,"denDF"]
    Fval <- object[effect.name,"F-value"]
    #etasq <- object$univariate.tests[effect.name,"SS"]/(object$univariate.tests[effect.name,"SS"] + object$univariate.tests[effect.name,"Error SS"])
    # generate latex output
    if(pval < .001) ptext <- "$p < .001$" else {
      if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
        ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
      }
    }
    txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
    #if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
    txt <- paste(txt,ptext,sep="")
    txt
  }
}

texreport.anova <- function(object,effect.name,effect.size=FALSE) {
  if(effect.name %in% rownames(object)) {
    pval <- object[effect.name,"Pr(>F)"]
    df1 <- object[effect.name,"NumDF"]
    df2 <- object[effect.name,"DenDF"]
    Fval <- object[effect.name,"F.value"]
    #etasq <- object$univariate.tests[effect.name,"SS"]/(object$univariate.tests[effect.name,"SS"] + object$univariate.tests[effect.name,"Error SS"])
    # generate latex output
    if(pval < .001) ptext <- "$p < .001$" else {
      if(pval < .05) ptext <- paste("$p = .",strsplit(as.character(round(pval,3)),".",fixed=TRUE)[[1]][2],"$",sep="") else {
        ptext <- paste("$p = .",strsplit(as.character(round(pval,2)),".",fixed=TRUE)[[1]][2],"$",sep="")
      }
    }
    txt <- paste("$F(",round(df1,2),",",round(df2,2),") = ",round(Fval,2),"$, ",sep="")
    #if(effect.size) txt <- paste(txt,"$\\eta^2_\\text{p} = ",round(etasq,2),"$, ",sep="")
    txt <- paste(txt,ptext,sep="")
    txt
  }
}

```

```{r read-data,warning=FALSE}
# source("R/readData.R")
# demo_exit <-  read.csv("data/demographics_and_exit_clean.csv")
# trial_data <- read.csv("data/trial_data_clean.csv")
# trial_data$trial <- 1:100
```

```{r, include=FALSE} 
options(tinytex.verbose = TRUE)
```

# Introduction
Being able to transfer previously acquired knowledge to a new domain is one of the hallmarks of human intelligence (Taylor et al., 2008). Humans are naturally endowed with the ability to extract relevant features from a situation, identify the presence of these features in a novel setting and use previously acquired knowledge to adapt to previously unseen challenges using acquired knowledge. More formally, Perkins (1992) defines transfer of learning as the application of skills, knowledge, and/or attitudes that were learned in one situation to another learning situation. This typically human skill has so far eluded modern ai agents. Deep neural networks for instance can do very well on image recognition tasks and can even reach super-human performance levels on video and strategic board games. Yet they struggle to learn as fast or as efficiently as humans do, and more importantly they have a very limited ability to generalize and transfer knowledge to new domains. Lake et al.(2017) argue that human learning transfer abilities take advantage of important cognitive building blocks such as a deeper representation of concepts underlying tasks and compositionally structured causal models of the environment.

In this paper, we are specifically interested in the way in which people build and use models of their opponent to facilitate learning transfer, when engaged in situations involving an interaction with strategic considerations. These situations arise frequently such as in negotiations, auctions, strategic planning and all other domains in which theory of mind abilities (Premack & Woodruff, 1978) play a role in determining human behaviour.

@@

Early literature on learning transfer in games has mostly focused on measuring the proportion of people who play normatively optimal (Nash Equilibria) or salient actions (e.g Risk Dominance) in later games, having had experience with a similar game environment previously. For instance, Ho et al. (1998) measure transfer as the proportion of players who choose the Nash Equilibrium in later p-beauty contest games, after traingn on similar games. They find there is no evidence of immediate transfer (Nash equilibrium play in the first round of the new game) but positive structural learning transfer as shown by the faster convergence to equilibrium play by experienced vs non experienced players. Camerer & Knez (2000) test learning transfer in players exposed to two games with multiple equilibria sequentially and explore the ability of players to conordinate their actions to choose a particular equilibrium in subsequent games having reached it in prior ones. They distinguish between games that are similar in a purely descriptive way, meaning similar choice labels, identity of players, format and number of action choices; and games that are similar in a strategic sense, meaning similar payoffs from combination of actions, identical equilibrium properties or significant social characteristics of payoffs such as possibility of punishment, need for fairness and cooperative vs competitive settings. They find that transfer of learning (successful coordination) occurs more readily in the presence of both descriptive and strategic similarity. If the games were only strategically similar, then the transfer was much weaker.

Juvina et al.(2014)  made a similar distinction between what they deemed surface and deep similarities and find that both contribute to positive learning transfer. However, they show that surface similarity is not necessary for deep transfer and can either aid or block this type of transfer depending on whether it leads to congruent or incongruent actions in later games. In a series of experiments using economic signalling games, Cooper & Kagel (2003, 2004, 2008) find that participants who have learned to play according to Nash Equilibrium in one game can transfer this to subsequent games, even though the actions consistent with Nash Equilibrium in later games are different.  They show that this transfer is driven by the emergence of sophisticated players who are able to represent the strategic implications of their actions and reason about the consequences of changed opponent payoffs.

@@

In most of the experimental paradigms we saw, participants were playing other human players either one on one or as a group. In this case, players are learning about the game but so are their opponents. While the approach of matching groups of human players repeatedly to play economic games has ecological validity in recreating environments where social decision making can be studied, it complicates the detection and modelling of strategies used by participants. It is harder to focus on an individual and how her strategies are changing and adapting to the opponent's play if we cannot experimentally control the behaviour of the opponent. It is also difficult to disentangle the process of learning about the opponent from that of learning about the game structure and payoffs. One potential solution is to replace the human opponent with a computerised agent, allowing researchers to manipulate the opponent's behaviour and observe how the human participants adapt their behaviour. For instance, Spiliopoulos (2013) made humans play constant sum games against 3 computer opponents programmed to take advantage of known patterns in human play such as imperfect randomization and heuristics use and found that human participants do adapt to the opponent they are facing. Shachat & Swarthout (2004) made human participants face computer opponents playing various mixed strategies in a zero-sum asymmetric matching pennies game. They found that the players changed their strategies towards exploiting the deviations from the MSNE, and that this exploitation was very likely if the deviation from the MSNE play was high.

One of the major limitations in studies of how humans adapt to computerised opponents is that they have mostly looked at the ability of players to detect and exploit action-based learning rules. They either used deterministic strategies that did not consider the human participant's play, as in the case of pre-determined mixed strategy algorithms, or strategies that only depended on the player's patterns of past plays. These strategies did not necessarily take into account the computer's own past plays, or the degree of strategic reasoning of the human player about the computer. Their play was not "human-like" in the sense that humans are not very good at playing specific mixed strategies with any precision of at detecting patterns from long sequences of past play due to cognitive constraints. It is therefore important to have agents that "play like humans", and one way of achieving that is to embed theses agents with human-like theory of mind abilities based on limited steps of recursive reasoning. Simon (1972) explains that humans have limited cognitive capacities and as such cannot be expected to solve computationally intractable problems such as finding Nash equilibria. Instead, they will try to 'satisfice' by choosing a strategy that is adequate in a simplified model of the environment, rather than an optimal one. This concept finds its natural application in 'level-k' theory, first adopted by Stahl & Wilson (1995). It posits that deviations from Nash equilibrium solutions in simple games are explained by the fact that humans have a heterogenous degree of strategic sophistication. At the bottom of the ladder, level-0 players are non-strategic and play either randomly or use a salient strategy in the game environment (Arad & Rubinstein, 2012). Level-1 players are next up the ladder of strategic sophistication and will assume all their opponents belong to the level-0 category and as such will best respond to them given this assumption. Likewise, a level-2 player will choose actions that are the best response given the belief that all opponents are exactly one level below and so on.

@@

# Methods

@@

## first experiment:

@@

Participants were informed they would play three different games against the same computer opponent. Each participant plays all three games consecutively and in the same order described above. Participants were told that the opponent cannot cheat and will choose its actions simultaneously without knowledge of the participant's choice. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. The overall score of the players was translated into the bonus by making each point worth $£$0.02. This bonus is significant as players could increase the total payoff from the experiment by up to 60% assuming they'd won all rounds against the computer opponent.

## Second Experiment

In the second experiment, participants each played 3 games sequentially against both level-1 and level-2 computer opponents, rather than just one like in the first experiment. The three games were Rock-Paper-Scissors, Fire-Water-Grass, and the penalty shootout game. The first two games were identical to the ones used in the first experiment. In the final game (shootout) the participants played the role of the player shooting a football (soccer) penalty, and the AI opponent was the goalkeeper. Players had the choice between three actions, like in the first two games: Shooting the football to the left, right or centre of the goal. If the player shoots in a direction different from that of where the goalkeeper dives, they win the round and the goalkeeper loses. Otherwise, the goalkeeper catches the ball and the player loses the round. There is no possibility of ties in this game. Figure XXXXXXXXX shows a snapshot of play in the shootout game. What makes this game different however is that there are two ways to beat the opponent in each round: if we think the opponent is going to choose ''right'' in the next round, we can win by both choosing ''left'' and ''center''. A level-1 player (thinks that his opponent will repeat his last action) has two ways to win the next round. As such, we have programmed the level-2 computer program to choose randomly between the two possibilities that a level-1 player may choose.

@@

```{r feedback_rps, fig.cap = "Screenshot of the feedback at the end of a round of Rock-Paper-Scissors", fig.align = "center"}

# knitr::include_graphics("../Report/images/feedback_rps.png", dpi = 108)

```

@@

<!-- Figure\ \@ref(fig:feedback_rps) -->

@@

# Results

## First Experiment

@@

To analyse within and between game learning, we used a 2 (condition: level-1, level-2) by 3 (game: RPS, FWG, NUMBERS) by 2 (block: first half, second half) repeated measures ANOVA with the first factor varying between participants. As the number of participants facing level-2 agents was slightly higher than those facing level-1 agents (28 vs 24), the ANOVA is unbalanced. We therefore used the “afex” (Singmann & al., 2016) package in R, as it allows us to handle unbalanced designs and produces type III sum of squares by default.

There was a main effect of Game ( F(2,100) = 8.54, ges  = 0.05,  p = .0005), showing that average scores varied significantly over the games. Post-hoc pairwise comparisons showed that performance in the FWG game was significantly higher than in the RPS game ( t(100) =3.78, p = 0.0008 ), and the performance in NUMBERS was significantly lower than FWG game ( t(100) = -3.32 , p = 0.0024). The score in RPS was not significantly different from the score in NUMBERS however ( t(100) = 0.45 , p = 0.65). The main effect of Block ( F(1,50) = 22.51 , p < .001, ges = 0.03) shows that the average score in the first half of games (M = 0.29)  was significantly lower than in the second half  of the games played (M = 0.40), which translates to within-game learning : as players get more interaction with the agent, they learn how to win more often. The main effect of Condition (F(1,50) = 5.44, p = .024, ges = 0.05) indicates that scores were higher against the level-1 player (M = 0.43) than against the level-2 player (M = 0.27). This means that it was harder for participants, on average, to learn the strategy of the more sophisticated opponent (level-2) compared to that of the comparatively less sophisticated agent (level-1). Finally, the analysis showed a significant interaction effects of block by game ( F(2,100) = 6.92 , p = .002, ges = 0.02), indicating that within-game learning differed between the games, as well as a three-way interaction between condition, game, and block ( F(2,100) = 3.88 , p = .023, ges = 0.01).

### Learning transfer

As a measure for learning transfer, we focus on participants’ scores in the earliest rounds, excluding the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response. A group of players with no experience of the game are expected to have scores not significantly different from 0. Any significantly positive group average scores would therefore reflect prior learning from past experiences. In figure XXX, the average score across participants by game for rounds 2-6 are plotted. Scores are also averaged across levels of condition. We test the average scores for each game against a hypothesised value of 0 for a non-experienced player using parametric one sample t-tests. As expected for the initial RPS game, the average score is not significantly different from 0 ( t(148.85) = 1.04 , p = 0.89). In FWG, the score is significantly higher than 0 ( t(148.85) = 4.58 , p < 0.0001). Unexpectedly, this is also the case for NUMBERS ( t(148.85) = 3.00, p = 0.0092).

### Moderation of learning transfer by type of opponent and game similarity

@@

## Second Experiment

>>>>

\PassOptionsToPackage{unicode=true}{hyperref} % options for packages loaded elsewhere
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames*,x11names*}{xcolor}
%
\documentclass[,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provides euro and other symbols
\else % if luatex or xelatex
  \usepackage{unicode-math}
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage[]{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\usepackage{xcolor}
\usepackage{hyperref}
\hypersetup{
            pdftitle={Transfer of Learned Opponent Models in Zero Sum Games},
            pdfauthor={Ismail Guennouni~\& Maarten Speekenbrink},
            pdfkeywords={keywords},
            colorlinks=true,
            linkcolor=blue,
            filecolor=Maroon,
            citecolor=Blue,
            urlcolor=Blue,
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

% set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\shorttitle{draft}
\affiliation{
\vspace{0.5cm}
\textsuperscript{1} Department of Experimental Psychology, University College London}
\keywords{keywords\newline\indent Word count: X}
\usepackage{csquotes}
\usepackage{upgreek}
\captionsetup{font=singlespacing,justification=justified}

\usepackage{longtable}
\usepackage{lscape}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage[flushleft]{threeparttable}
\usepackage{threeparttablex}

\newenvironment{lltable}{\begin{landscape}\begin{center}\begin{ThreePartTable}}{\end{ThreePartTable}\end{center}\end{landscape}}

\makeatletter
\newcommand\LastLTentrywidth{1em}
\newlength\longtablewidth
\setlength{\longtablewidth}{1in}
\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}

\usepackage{lineno}

\linenumbers
\usepackage[document]{ragged2e}

\title{Transfer of Learned Opponent Models in Zero Sum Games}
\author{Ismail Guennouni\textsuperscript{1}~\& Maarten Speekenbrink\textsuperscript{1}}
\date{}

\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

@@

Correspondence concerning this article should be addressed to Ismail Guennouni, Department of Experimental Psychology, University College London, 26 Bedford Way, London WC1H 0AP, United Kingdom. E-mail: \href{mailto:i.guennouni.17@ucl.ac.uk}{\nolinkurl{i.guennouni.17@ucl.ac.uk}}}

\abstract{
Enter abstract here. Each new line herein must be indented, like this line.

}

\begin{document}
\maketitle

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Being able to transfer previously acquired knowledge to a new domain is one of the hallmarks of human intelligence (Taylor et al., 2008). Humans are naturally endowed with the ability to extract relevant features from a situation, identify the presence of these features in a novel setting and use previously acquired knowledge to adapt to previously unseen challenges using acquired knowledge. More formally, Perkins (1992) defines transfer of learning as the application of skills, knowledge, and/or attitudes that were learned in one situation to another learning situation. This typically human skill has so far eluded modern ai agents. Deep neural networks for instance can do very well on image recognition tasks and can even reach super-human performance levels on video and strategic board games. Yet they struggle to learn as fast or as efficiently as humans do, and more importantly they have a very limited ability to generalize and transfer knowledge to new domains. Lake et al.(2017) argue that human learning transfer abilities take advantage of important cognitive building blocks such as a deeper representation of concepts underlying tasks and compositionally structured causal models of the environment.

In this paper, we are specifically interested in the way in which people build and use models of their opponent to facilitate learning transfer, when engaged in situations involving an interaction with strategic considerations. These situations arise frequently such as in negotiations, auctions, strategic planning and all other domains in which theory of mind abilities (Premack \& Woodruff, 1978) play a role in determining human behaviour.

@@

Early literature on learning transfer in games has mostly focused on measuring the proportion of people who play normatively optimal (Nash Equilibria) or salient actions (e.g Risk Dominance) in later games, having had experience with a similar game environment previously. For instance, Ho et al. (1998) measure transfer as the proportion of players who choose the Nash Equilibrium in later p-beauty contest games, after traingn on similar games. They find there is no evidence of immediate transfer (Nash equilibrium play in the first round of the new game) but positive structural learning transfer as shown by the faster convergence to equilibrium play by experienced vs non experienced players. Camerer \& Knez (2000) test learning transfer in players exposed to two games with multiple equilibria sequentially and explore the ability of players to conordinate their actions to choose a particular equilibrium in subsequent games having reached it in prior ones. They distinguish between games that are similar in a purely descriptive way, meaning similar choice labels, identity of players, format and number of action choices; and games that are similar in a strategic sense, meaning similar payoffs from combination of actions, identical equilibrium properties or significant social characteristics of payoffs such as possibility of punishment, need for fairness and cooperative vs competitive settings. They find that transfer of learning (successful coordination) occurs more readily in the presence of both descriptive and strategic similarity. If the games were only strategically similar, then the transfer was much weaker.

Juvina et al.(2014) made a similar distinction between what they deemed surface and deep similarities and find that both contribute to positive learning transfer. However, they show that surface similarity is not necessary for deep transfer and can either aid or block this type of transfer depending on whether it leads to congruent or incongruent actions in later games. In a series of experiments using economic signalling games, Cooper \& Kagel (2003, 2004, 2008) find that participants who have learned to play according to Nash Equilibrium in one game can transfer this to subsequent games, even though the actions consistent with Nash Equilibrium in later games are different. They show that this transfer is driven by the emergence of sophisticated players who are able to represent the strategic implications of their actions and reason about the consequences of changed opponent payoffs.

@@

In most of the experimental paradigms we saw, participants were playing other human players either one on one or as a group. In this case, players are learning about the game but so are their opponents. While the approach of matching groups of human players repeatedly to play economic games has ecological validity in recreating environments where social decision making can be studied, it complicates the detection and modelling of strategies used by participants. It is harder to focus on an individual and how her strategies are changing and adapting to the opponent's play if we cannot experimentally control the behaviour of the opponent. It is also difficult to disentangle the process of learning about the opponent from that of learning about the game structure and payoffs. One potential solution is to replace the human opponent with a computerised agent, allowing researchers to manipulate the opponent's behaviour and observe how the human participants adapt their behaviour. For instance, Spiliopoulos (2013) made humans play constant sum games against 3 computer opponents programmed to take advantage of known patterns in human play such as imperfect randomization and heuristics use and found that human participants do adapt to the opponent they are facing. Shachat \& Swarthout (2004) made human participants face computer opponents playing various mixed strategies in a zero-sum asymmetric matching pennies game. They found that the players changed their strategies towards exploiting the deviations from the MSNE, and that this exploitation was very likely if the deviation from the MSNE play was high.

One of the major limitations in studies of how humans adapt to computerised opponents is that they have mostly looked at the ability of players to detect and exploit action-based learning rules. They either used deterministic strategies that did not consider the human participant's play, as in the case of pre-determined mixed strategy algorithms, or strategies that only depended on the player's patterns of past plays. These strategies did not necessarily take into account the computer's own past plays, or the degree of strategic reasoning of the human player about the computer. Their play was not \enquote{human-like} in the sense that humans are not very good at playing specific mixed strategies with any precision of at detecting patterns from long sequences of past play due to cognitive constraints. It is therefore important to have agents that \enquote{play like humans}, and one way of achieving that is to embed theses agents with human-like theory of mind abilities based on limited steps of recursive reasoning. Simon (1972) explains that humans have limited cognitive capacities and as such cannot be expected to solve computationally intractable problems such as finding Nash equilibria. Instead, they will try to \enquote{satisfice} by choosing a strategy that is adequate in a simplified model of the environment, rather than an optimal one. This concept finds its natural application in \enquote{level-k} theory, first adopted by Stahl \& Wilson (1995). It posits that deviations from Nash equilibrium solutions in simple games are explained by the fact that humans have a heterogenous degree of strategic sophistication. At the bottom of the ladder, level-0 players are non-strategic and play either randomly or use a salient strategy in the game environment (Arad \& Rubinstein, 2012). Level-1 players are next up the ladder of strategic sophistication and will assume all their opponents belong to the level-0 category and as such will best respond to them given this assumption. Likewise, a level-2 player will choose actions that are the best response given the belief that all opponents are exactly one level below and so on.

@@

\hypertarget{methods}{%
\section{Methods}\label{methods}}

@@

\hypertarget{first-experiment}{%
\subsection{first experiment:}\label{first-experiment}}

@@

Participants were informed they would play three different games against the same computer opponent. Each participant plays all three games consecutively and in the same order described above. Participants were told that the opponent cannot cheat and will choose its actions simultaneously without knowledge of the participant's choice. A total of 50 rounds of each game was played with the player's score displayed at the end of each game. The score was calculated as the number of wins minus the number of losses. Ties did not affect the score. In order to incentivise the participants to maximise the number of wins against the opponents, players were paid a bonus at the end of the experiment that was proportional to their final score. The overall score of the players was translated into the bonus by making each point worth \$£\$0.02. This bonus is significant as players could increase the total payoff from the experiment by up to 60\% assuming they'd won all rounds against the computer opponent.

\hypertarget{second-experiment}{%
\subsection{Second Experiment}\label{second-experiment}}

In the second experiment, participants each played 3 games sequentially against both level-1 and level-2 computer opponents, rather than just one like in the first experiment. The three games were Rock-Paper-Scissors, Fire-Water-Grass, and the penalty shootout game. The first two games were identical to the ones used in the first experiment. In the final game (shootout) the participants played the role of the player shooting a football (soccer) penalty, and the AI opponent was the goalkeeper. Players had the choice between three actions, like in the first two games: Shooting the football to the left, right or centre of the goal. If the player shoots in a direction different from that of where the goalkeeper dives, they win the round and the goalkeeper loses. Otherwise, the goalkeeper catches the ball and the player loses the round. There is no possibility of ties in this game. Figure XXXXXXXXX shows a snapshot of play in the shootout game. What makes this game different however is that there are two ways to beat the opponent in each round: if we think the opponent is going to choose \enquote{'right}' in the next round, we can win by both choosing \enquote{'left}' and \enquote{'center}'. A level-1 player (thinks that his opponent will repeat his last action) has two ways to win the next round. As such, we have programmed the level-2 computer program to choose randomly between the two possibilities that a level-1 player may choose.

@@

@@

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{first-experiment-1}{%
\subsection{First Experiment}\label{first-experiment-1}}

@@

To analyse within and between game learning, we used a 2 (condition: level-1, level-2) by 3 (game: RPS, FWG, NUMBERS) by 2 (block: first half, second half) repeated measures ANOVA with the first factor varying between participants. As the number of participants facing level-2 agents was slightly higher than those facing level-1 agents (28 vs 24), the ANOVA is unbalanced. We therefore used the \enquote{afex} (Singmann \& al., 2016) package in R, as it allows us to handle unbalanced designs and produces type III sum of squares by default.

There was a main effect of Game ( F(2,100) = 8.54, ges = 0.05, p = .0005), showing that average scores varied significantly over the games. Post-hoc pairwise comparisons showed that performance in the FWG game was significantly higher than in the RPS game ( t(100) =3.78, p = 0.0008 ), and the performance in NUMBERS was significantly lower than FWG game ( t(100) = -3.32 , p = 0.0024). The score in RPS was not significantly different from the score in NUMBERS however ( t(100) = 0.45 , p = 0.65). The main effect of Block ( F(1,50) = 22.51 , p \textless{} .001, ges = 0.03) shows that the average score in the first half of games (M = 0.29) was significantly lower than in the second half of the games played (M = 0.40), which translates to within-game learning : as players get more interaction with the agent, they learn how to win more often. The main effect of Condition (F(1,50) = 5.44, p = .024, ges = 0.05) indicates that scores were higher against the level-1 player (M = 0.43) than against the level-2 player (M = 0.27). This means that it was harder for participants, on average, to learn the strategy of the more sophisticated opponent (level-2) compared to that of the comparatively less sophisticated agent (level-1). Finally, the analysis showed a significant interaction effects of block by game ( F(2,100) = 6.92 , p = .002, ges = 0.02), indicating that within-game learning differed between the games, as well as a three-way interaction between condition, game, and block ( F(2,100) = 3.88 , p = .023, ges = 0.01).

\hypertarget{learning-transfer}{%
\subsubsection{Learning transfer}\label{learning-transfer}}

As a measure for learning transfer, we focus on participants' scores in the earliest rounds, excluding the very first round for which the agent is programmed to play randomly as it has no data on prior rounds on which to build its response. A group of players with no experience of the game are expected to have scores not significantly different from 0. Any significantly positive group average scores would therefore reflect prior learning from past experiences. In figure XXX, the average score across participants by game for rounds 2-6 are plotted. Scores are also averaged across levels of condition. We test the average scores for each game against a hypothesised value of 0 for a non-experienced player using parametric one sample t-tests. As expected for the initial RPS game, the average score is not significantly different from 0 ( t(148.85) = 1.04 , p = 0.89). In FWG, the score is significantly higher than 0 ( t(148.85) = 4.58 , p \textless{} 0.0001). Unexpectedly, this is also the case for NUMBERS ( t(148.85) = 3.00, p = 0.0092).

\hypertarget{moderation-of-learning-transfer-by-type-of-opponent-and-game-similarity}{%
\subsubsection{Moderation of learning transfer by type of opponent and game similarity}\label{moderation-of-learning-transfer-by-type-of-opponent-and-game-similarity}}

@@

\hypertarget{second-experiment-1}{%
\subsection{Second Experiment}\label{second-experiment-1}}
